Reanalysis Step 2: Multiple Imputation

### 2025-08-30 reanalysis-2-multiple_imputation.qmd
consistent person dr

  # Copy person_dr_max back to person_dr for each eye
  # This ensures DR severity is consistent within patients
  imp_data <- imp_data |>
    mutate(
      # Convert person_dr_max back to numeric for person_dr
      person_dr = as.numeric(as.character(factor(person_dr_max, 
                                                 labels = c("0", "1", "2"))))
    )

@@

```{r add-back-vars}
if(length(vars_to_add) > 0) {
  # Add these variables back to each imputed dataset using row_id for exact matching
  imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
    # Store original row count
    orig_rows <- nrow(imp_data)

    # Join using row_id for exact 1:1 matching
    imp_data <- imp_data |>
      left_join(
        df_original_full |> 
          dplyr::select(row_id, dplyr::all_of(vars_to_add)),
        by = "row_id"
      )

>>>>

  # Copy person_dr_max back to person_dr for each eye
  # This ensures DR severity is consistent within patients
  # Handle the factor conversion correctly
  imp_data <- imp_data |>
    mutate(
      # person_dr_max is an ordered factor with levels "No_DR", "NPDR", "PDR"
      # Convert to numeric 0, 1, 2
      person_dr = case_when(
        person_dr_max == "No_DR" ~ 0,
        person_dr_max == "NPDR" ~ 1,
        person_dr_max == "PDR" ~ 2,
        TRUE ~ NA_real_
      )
    ) |>
    select(-person_dr_max)  # Remove person_dr_max after using it

@@

```{r add-back-vars}
if(length(vars_to_add) > 0) {
  # IMPORTANT: Exclude No_DR, NPDR, PDR from vars_to_add to prevent overwriting person_dr
  # These would cause create_dr_severity_variables to recalculate person_dr from eye-level data
  original_vars_to_add <- vars_to_add
  vars_to_add <- setdiff(vars_to_add, c("No_DR", "NPDR", "PDR"))

  if (length(original_vars_to_add) != length(vars_to_add)) {
    log_message("Excluded No_DR/NPDR/PDR from variables to add back to preserve person_dr consistency")
  }

  # Add these variables back to each imputed dataset using row_id for exact matching
  imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
    # Store original row count
    orig_rows <- nrow(imp_data)

    # Store the consistent person_dr we just created
    person_dr_consistent <- imp_data$person_dr

    # Join using row_id for exact 1:1 matching
    imp_data <- imp_data |>
      left_join(
        df_original_full |> 
          dplyr::select(row_id, dplyr::all_of(vars_to_add)),
        by = "row_id"
      ) |>
      # Ensure person_dr remains consistent (in case it somehow got overwritten)
      mutate(person_dr = person_dr_consistent)
